---
title: "705604096_stats101a_hw2"
author: "Jade Gregory"
date: "2023-04-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
a)

```{r}
n <- 30
mean <- 23606
sd <- 24757
lb <- mean - abs(qt(.025, 29) * sd / sqrt(n))
ub <- mean + abs(qt(.025, 29) * sd / sqrt(n))
lb
ub
```
The 95% confidence interval is (14361.58, 32850.42)

b)
We have to assume the population distribution is Normal or our sample size is sufficiently large enough to provide us with a good approximation.

c)
No, because a different sample will produce different sample statistics used to compute the confidence interval. 

d)
```{r}
a <- 25000
m <- 23606
sd <- 24757
n <- 30
z <- (m - a) / (sd / sqrt(n))
2 * pnorm(-abs(z))
```
Therefore, since our p-value is 0.757772 which is greater than our significance level of 0.05, we fail to reject the null hypothesis. So, the mean income is $25,000.

e)
The smallest significance level we could have used so that we reject the null hypothesis would be 0.76.

## Question 2
a)
```{r}
n <- 100
mean <- 23606
sd <- 24757
lb <- mean - abs(qt(.025, 99) * sd / sqrt(n))
ub <- mean + abs(qt(.025, 99) * sd / sqrt(n))
lb
ub
```
The width of the interval decreases as the sample size increases.

b)
The width of the interval decreases as our confidence level decreases.

## Question 3
```{r}
cdc <- read.csv('cdc copy.csv')
t.test(cdc$weight ~ cdc$exerany, alternative = "two.sided", conf.level = 0.95)
```
a)
Let $$\bar{x}_e$$ represent the mean of desired weight for people who exercise and let $$\bar{x}_n$$ represent the mean of desired weight for people who do not exercise.
$$
H_0 : \bar{x}_e = \bar{x}_n \\
H_a : \bar{x}_e \neq \bar{x}_n
$$
b)
The value of our test statistic is 3.6842.

c)
The p-value of our test statistic is 0.0002309

d)
With a significance level of 0.05, we reject our null hypothesis. 

e)
No, this is not the correct definition of the p-value. The p-value measures how likely it is that any observed difference between the null hypothesis and alternative hypothesis is due to chance. 

f)
The significance level is a measure of how small we want the probability of rejecting the null hypothesis when it is actually true to be. Using the p-value, it tests the probability of observing extreme values by chance. 