---
title: "705604096_stats101a_hw1"
author: "Jade Gregory"
date: "2023-04-05"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(stringr)
library(tinytex)
library(readr)
```

Jade Gregory
705604096

## Question 1
```{r}
df <- read_csv('chicagotaxiraw.csv')
df
```
```{r}
df_new <- df[,c(1:6,11:17)]
df_new
```
```{r}
df_final <- rename(df_new, c(trip_id = 'Trip ID', taxi_id = 'Taxi ID', 
                             trip_start_timestamp = 'Trip Start Timestamp',
                             trip_end_timestamp = 'Trip End Timestamp',
                             trip_seconds = 'Trip Seconds', trip_miles = 'Trip Miles',
                             fare = Fare, tips = Tips, tolls = Tolls, extras = Extras,
                             trip_total = 'Trip Total', payment_type = 'Payment Type',
                             company = Company))
df_final
```
## Question 2

```{r}
min(df_final$trip_start_timestamp)
max(df_final$trip_end_timestamp, na.rm = TRUE)
```

This data ranges from 1/1/2013 to 5/31/2017. The date of the first pickup is 1/1/2013 and the date of the last
drop-off is 5/31/2017.

## Question 3

```{r}
df_final$weekday <- wday(df_final$trip_start_timestamp, week_start = 1)
df_final
```
```{r}
df_mon <- filter(df_final, weekday == 1)
length(df_mon$weekday)
```
```{r}
df_tue <- filter(df_final, weekday == 2)
length(df_tue$weekday)
```
```{r}
df_wed <- filter(df_final, weekday == 3)
length(df_wed$weekday)
```
```{r}
df_thur <- filter(df_final, weekday == 4)
length(df_thur$weekday)
```
```{r}
df_fri <- filter(df_final, weekday == 5)
length(df_fri$weekday)
```
```{r}
df_sat <- filter(df_final, weekday == 6)
length(df_sat$weekday)
```
```{r}
df_sun <- filter(df_final, weekday == 7)
length(df_sun$weekday)
```

From our output, we can see that Friday is the day of the week to have the highest number of trips begin
with a total of 16767 trips began throughout this data set. Monday is the day of the week with the least amount 
of trips started with a total of 12406 trips started throughout this data set.

## Question 4

```{r}
tips <- df_final$tips
tips <- gsub("\\$", "", tips)
tips <- as.numeric(tips)
mean(tips, na.rm = TRUE)
median(tips)
```
The typical value for the tip is $0. The mean of the tip amount is better to use since it is an
average of how much people tip the taxis.



