---
title: "705604096_stats101a_hw3"
author: "Jade Gregory"
date: "2023-04-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

## Question 1
A)
a)
```{r}
slr <- function(beta_0, beta_1, sigma, x, rns){
  set.seed(rns)
  eps <- rnorm(length(x), 0, sigma)
  beta_0 + (beta_1 * x) + eps
}
```

b)
```{r}
x <- rep(1:10, by = .1, 4)
y <- slr(1, 3, 3, x, 123)
plot(x, y, main = "Scatterplot")
```

B)
```{r}
cor(x, y)
```
The correlation coefficient is 0.9529631

C)
```{r}
x2 <- rep(1:10, by = .1, 4)
y2 <- slr(1, 30, 3, x2, 123)
cor(x2, y2)
```
With parameters $$\beta_0 = 1$$ $$ \beta_1 = 30 $$ $$ \sigma = 3$$ we get a correlation of 0.9995272

## Question 2
```{r}
my_data <- read_csv('armspans2022_gender.csv')
my_data['compmother'] <- tolower(my_data$compmother) 
my_data['compfather'] <- tolower(my_data$compfather) 
my_data$compmother <- str_replace_all(my_data$compmother, c(" " = "_", "'" =""))
my_data$compfather <- str_replace_all(my_data$compfather, c(" " = "_", "'" =""))
my_data
```
When cleaning this data, I decided to make all of the elements in the columns lower case to make it easier to manipulate in the future. I also replaced spaces in between words with underscores for the same reason.

a)
```{r}
plot(armspan ~ height, data = my_data )
```

We can observe a positive linear trend in this scatter plot. We can describe the strength as in between moderate and strong between the observations, since there is a clear linear pattern that is noticeable. There are not any unusual features we can discern from this scatter plot, except for the lowest value being farther from the bulk of the data but still in line with our linear pattern.

b)
```{r}
plot(armspan ~ height, data = my_data )
my_lm <- lm(armspan ~ height, data = my_data)
summary(my_lm)
abline(my_lm)
```

The equation for our estimated line is $$\hat{y}_i = -29.63530 + 1.42459x_i$$

c)
```{r}
my_func <- function(x){
  -29.6353 + 1.42459 * x
}
my_func(63)
residual <- 61 - my_func(63)
residual
```

Based on our model, the predicted arm span for my height is 60.11387 inches. The residual is 0.88613 inches.

d)
```{r}
my_func(76)
residual2 <- 79 - my_func(76)
residual2
```

No, this does not seem unusual as the residual is relatively small.

e)
```{r}
res <- resid(my_lm)
plot(fitted(my_lm), res)
abline(0,0)
```

Because there does not seem to be any pattern in the residual points with respect to the line at x = 0, we can conclude that the linear model is the right fit for this data. We can also note that there is a fan shape that appears in our residual plot, with exception to the outlier as previously mentioned, that would indicate that the standard deviation increases as x increases.

## Question 3
a)
```{r}
quad_func <- function(a, b, c, sigma, x = rep(1:10,by=.1,4), random.seed){
  set.seed(random.seed)
  a + (b * x) + (c * (x^2)) + rnorm(length(x), 0, sigma)
}
x <- rep(1:10,by=.1,4)
y <- quad_func(1, 3, 1, 3, x, 123)
plot(x, y)
```
b)
```{r}
plot(x, y)
my_lm3 <- lm(y ~ x)
abline(my_lm3)
```
```{r}
res3 <- resid(my_lm3)
plot(fitted(my_lm3), res3)
abline(0,0)
```

In this residual plot, we can observe a quadratic shape or a parabola appear.

c)
If the residual plot shows features like parabolic shapes, we can conclude that the trend is non-linear.

d)
```{r}
my_nl_func <- function(a, b, x, sigma, random.seed){
  set.seed(random.seed)
  a + (b * x) + rnorm(length(x), 0, (sigma * (x^2)))
}
x4 <- rep(1:10, by = .1, 4)
y4 <- my_nl_func(1, 200, x4, 5, 123)
plot(y4 ~ x4)
```
e)
```{r}
plot(x4, y4)
my_lm4 <- lm(y4 ~ x4)
abline(my_lm4)
```
```{r}
res4 <- resid(my_lm4)
plot(fitted(my_lm4), res4)
abline(0,0)
```

The cone shape of the residual plot indicates that the standard deviation increases as x increases, and therefore the constant standard deviation assumption is violated.

## Question 4
```{r}
atus <- read.csv('atus.csv')
atus1 <- subset(atus, homework > 0)
head(atus1)
```
a)
```{r}
plot(homework ~ sleep, data = atus1)
my_lm_2 <- lm(homework ~ sleep, data = atus1)
abline(my_lm_2)
```

The linear model is not an appropriate model to describe the relationship between sleep times and homework times.

b)
```{r}
res2 <- resid(my_lm_2)
plot(fitted(my_lm_2), res2)
abline(0,0)
```

From the residual plot, we can tell that this data is negatively correlated. We can also tell that since it is not evenly distributed, a linear model would not be the best model to approximate this data. 

## Question 5
a)
```{r}
t.test(atus1$household_chores ~ atus1$gender, alternative = "greater", conf.level = 0.95)
```
Let $$\bar{x}_f$$ represent the average time persons identifying as female spent doing chores
Let $$\bar{x}_m$$ represent the average time persons identifying as male spent doing chores
$$
H_0 : \bar{x}_f - \bar{x}_m = 0
$$
$$
H_a : \bar{x}_f - \bar{x}_m \neq 0
$$
The test statistic is 6.3978. the observed value of the statistic is 446.68. The p-value is 1.993e-10. With a 5% significance level, we reject the null hypothesis.

b)
We must assume the population distribution is normal or our sample size is sufficiently large to provide us with a good approximation. We can assume that these conditions are met because we have a large sample size observed in our data frame, which is sufficient.

